---
title: "HW6"
format: html
editor: visual
---

```{r}
library(purrr)
library(httr)
library(jsonlite)
library(dplyr)
library(lubridate)
```


## Task 1

1.  What is the purpose of the lapply() function? What is the equivalent purrr function?

> lapply takes a function and applies it to every element in a list. purrr has map to do the same thing.

2.  Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall") on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify method = "kendall" when calling lapply())

> lapply(my_list, cor, method = "kendall")

3.  What are two advantages of using purrr functions instead of the BaseR apply family?

> purrr maintains consistency with the ordering and purpose of its arguments and has type stability

4.  What is a side-effect function?

> It is a function that does something, but in the context of a functional-style programming chain, it doesn't modify the value that is being passed from function to function.

5.  Why can you name a variable sd in a function and not cause any issues with the sd function?

> This is due to R's scoping method where a variable is first searched for in the function environment before progressing outside to the global environment and beyond.

## Task 2

first write RMSE function

```{r}
getRMSE <- function(responses, predictions, ...){
  squared_diffs <- (responses - predictions) ^ 2
  return(sqrt(mean(squared_diffs, ...)))
}
```

now we'll create some test data for the function

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

and we'll test it several ways

```{r}
getRMSE(resp, pred)

resp[c(4, 82)] <- NA_real_

getRMSE(resp, pred)
getRMSE(resp, pred, na.rm = TRUE)
```

we repeat the process for MAE

```{r}
getMAE = function(responses, predictions, ...){
  absolute_diffs <- abs(responses - predictions)
  return(sqrt(mean(absolute_diffs, ...)))
}

set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

getMAE(resp, pred)

resp[c(13, 54)] <- NA_real_

getMAE(resp, pred)
getMAE(resp, pred, na.rm = TRUE)
```

now we combine into a wrapper function

```{r}
getPredictionMetrics <- function(responses, predictions, modes = c("RMSE", "MAE"), ...){
  
  if(!is.atomic(responses) || !is.numeric(responses) || !is.atomic(predictions) || !is.numeric(predictions)){
    stop("Either responses or predictions are not an atomic numeric vector")
  }
  
  results <- map(modes, function(mode){
    metric_function <- get(paste(c("get", mode), collapse = ""))
    return(metric_function(responses, predictions, ...))
  })
  
  names(results) <- modes
  
  return(results)
}
```

now we test as requested

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

getPredictionMetrics(resp, pred)
getPredictionMetrics(resp, pred, modes = c("MAE"))
getPredictionMetrics(resp, pred, modes = c("RMSE"))

resp[c(34, 67)] <- NA_real_

getPredictionMetrics(resp, pred)
getPredictionMetrics(resp, pred, na.rm = TRUE)
#had to comment out this line because the error was effective
#getPredictionMetrics(resp, c("Hello", "this", "is", "invalid", "data"))
```

## Task 3

first get and unpack API response

```{r}
full_resp <- GET("https://newsapi.org/v2/everything", query = list(q = "click to cancel", apiKey = "675b998446044db1b34cd9f57dcfa261"))

news_results <- full_resp$content |> 
  rawToChar() |> 
  fromJSON() |> 
  as_tibble() |> 
  pluck("articles")
```

now we package it as a function

```{r}
get_news_data <- function(search_terms, date_from, key){
  full_resp <- GET("https://newsapi.org/v2/everything", query = list(q = search_terms, from = date_from, apiKey = key))

return(full_resp$content |> 
  rawToChar() |> 
  fromJSON() |>  
  pluck("articles") |> 
  as_tibble())
}
```

last, we process

```{r}
ohtani_tbl <- get_news_data("Shohei Ohtani", date_from = "2024-10-01", key = "675b998446044db1b34cd9f57dcfa261")

judge_tbl <- get_news_data("Aaron Judge", date_from = "2024-10-01", key = "675b998446044db1b34cd9f57dcfa261")

ohtani_tbl |> 
  pluck("source", "name") |> 
  table()

judge_tbl |> 
  pluck("source", "name") |> 
  table()

ohtani_tbl <- ohtani_tbl |> 
  mutate(publishedAt = ymd_hms(publishedAt)) |> 
  arrange(publishedAt) |> 
  mutate(pub_diff = publishedAt - lag(publishedAt))

judge_tbl <- judge_tbl |> 
  mutate(publishedAt = ymd_hms(publishedAt)) |> 
  arrange(publishedAt) |> 
  mutate(pub_diff = publishedAt - lag(publishedAt))

ohtani_tbl |> 
  select(publishedAt, pub_diff) |> 
  map(\(x) list(mean = mean(x, na.rm = TRUE), sd = sd(x, na.rm = TRUE), median = median(x, na.rm = TRUE)))
```